Statistical Machine Learning Models on A Bank Marketing Dataset

### 2025-12-09 Slides.Rmd
final slides update

---
title: "Statistical Machine Learning Models on A Bank Marketing Dataset"
author: "Sandra, Tigist, Mussa Hassen, Waddah"
subtitle: 'Colorado State University, DSCI 445: Statistical Machine Learning, Dr.Andee
  Kaplan'
output:
  ioslides_presentation:
    widescreen: true
---

@@

## Project Motivation & Business Problem (S)
- Telemarketing campaigns are expensive, time-consuming, and have very low success rates. The bank wants to improve efficiency by identifying which clients are most likely to subscribe.
- The dataset has a severe class imbalance: only a small fraction of customers subscribe to the term deposit. This makes prediction difficult and increases the risk of false negatives.
- Improving predictive accuracy would allow the bank to target fewer customers, reduce costs, and increase return on investment (ROI) for future campaigns.

```{r, echo=FALSE}
library(ggplot2)

ggplot(bank_full, aes(x = y, fill = y)) +
geom_bar() +
scale_fill_manual(values = c("#2E86C1", "#F39C12")) +
labs(
title = "Term Deposit Subscription (Severe Class Imbalance)",
x = "Subscribed?",
y = "Count"
) +
theme_minimal(base_size = 14)

@@

## Methodology: Data Preparation (Slide 5) (s)

@@

## Boosting 2/3: Hyperparameter Tuning 
- We tuned the number of trees using OOB error estimation to prevent overfitting.
- The OOB method selected **`r best_trees`** trees as optimal.
- This choice balances model complexity and generalization.

>>>>

---
title: "Statistical Machine Learning Models on A Bank Marketing Dataset"
author: "Sandra Afrifa, Tigist Kefelew, Mussa Hassen, Waddah"
subtitle: 'Colorado State University, DSCI 445: Statistical Machine Learning, Dr.Andee
  Kaplan'
output:
  ioslides_presentation:
    widescreen: true
---

@@

## Project Motivation & Business Problem
- Telemarketing campaigns are expensive, time-consuming, and have very low success rates. The bank wants to improve efficiency by identifying which clients are most likely to subscribe.
- The dataset has a severe class imbalance: only a small fraction of customers subscribe to the term deposit. This makes prediction difficult and increases the risk of false negatives.
- Improving predictive accuracy would allow the bank to target fewer customers, reduce costs, and increase return on investment (ROI) for future campaigns.

@@

## Methodology: Data Preparation

@@

## Boosting 2/3: Hyperparameter Tuning 
- We tuned the number of trees using OOB error estimation to prevent overfitting.
- The OOB method selected 1100 trees as optimal.
- This choice balances model complexity and generalization.
